Reasoning Graphically about Quantum Computation

### 2008-03-04 papers/reasoning-graphically/paper.tex
...
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1399 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

Graphs with certain additional structure give a
representation for compact closed  categories; we now give an overview
of this construction.  The details omitted here can be found in
\cite{Duncan:thesis:2006}.   Pictorial representations are in
Fig.~\ref{fig:comcl-graphs}.

\begin{figure}
  We make the convention that domain of an arrow is at the top of the
  picture, and its codomain is at the bottom.
  \centering
  \[
  \begin{array}{lllll}
      \id{A \otimes B^*} = \inlinegraphic{3em}{comcl-id}
      &\qquad&
      d_A = \inlinegraphic{2em}{comcl-eta}
      &\qquad& 
      e_B = \inlinegraphic{2em}{comcl-epsilon} 
      \\\\
      f = \inlinegraphic{3em}{comcl-f} 
      & \qquad &
      \name{f} = \inlinegraphic{3em}{comcl-name-f}
      &\qquad &
      f^* =  \inlinegraphic{3em}{comcl-dual-f}
  \end{array}
  \]
  \caption{Compact Closed Structure as Graphs}
  \label{fig:comcl-graphs}
\end{figure}

@@

Hence any connected subgraph
involving nodes of only one colour, wherever it occurs in a graph, may
be collapsed to a single vertex labelled by a single value $\alpha$,
giving a ``spider'' as depicted below:
\[
\inlinegraphic{3em}{gen-spider}
\]
Conversely, a spider may be arbitrarily divided into subspiders,
provided the total in- and out-degree is preserved, along with and the
sum of the $\alpha $s.  Further, one can derive from the spider
theorem $n$-fold versions of many of the other rewrites in the system.
Informally, this can be expressed as the graph equation in
Figure~\ref{informal-spider-law-fig}.

Spiders offer a very intuitive way to manipulate graphs,
and are far more compact and convenient in calculations than the
graphs built up naively 
from the generators.  However, formalising spiders requires moving from
finite graphs, where each vertex has bounded degree, and which are
subject to a finite number of rewrite rules, to a system where nodes
may have arbitrarily many edges, and there are infinitely many rewrite
rules.  The desire to retain these intuitive reasoning methods
motivates the extension from concrete graphs to \emph{graph patterns}.
These will comprise the main subject of this paper.

@@

>>>>

Graphs with certain additional structure give a representation for
compact closed categories; we now give an overview of this
construction.  The details omitted here can be found in
\cite{Duncan:thesis:2006}.  Pictorial representations are in
Fig.~\ref{fig:comcl-graphs}. We make the convention that domain of an
arrow is at the top of the picture, and its codomain is at the bottom.

\begin{figure}[t]
  \centering
  \[
  \begin{array}{lllll}
      \id{A \otimes B^*} = \inlinegraphic{3em}{comcl-id}
      &\qquad&
      d_A = \inlinegraphic{2em}{comcl-eta}
      &\qquad& 
      e_B = \inlinegraphic{2em}{comcl-epsilon} 
      \\\\
      f = \inlinegraphic{3em}{comcl-f} 
      & \qquad &
      \name{f} = \inlinegraphic{3em}{comcl-name-f}
      &\qquad &
      f^* =  \inlinegraphic{3em}{comcl-dual-f}
  \end{array}
  \]
  \caption{Compact Closed Structure as Graphs. }
  \label{fig:comcl-graphs}
\end{figure}

@@

\begin{figure}[t]
\includegraphics{images/gen-spider} = \includegraphics{images/gen-spider}
\label{fig:informal-spider-law}\caption{An informal equation on graphs
  that expresses the Spider Theorem. }
\end{figure}

Hence any connected subgraph involving nodes of only one colour,
wherever it occurs in a graph, may be collapsed to a single vertex
labelled by a single value $\alpha$, giving a ``spider''. Informally,
this can be depicted graphically as the equation in
Figure~\ref{fig:informal-spider-law}. Conversely, a spider may be
arbitrarily divided into subspiders, provided the total in- and
out-degree is preserved, along with and the sum of the $\alpha $s.
Further, one can derive, from the spider theorem, $n$-fold versions of
many of the other equations.

Spiders offer a very intuitive way to manipulate graphs, and are far
more compact and convenient in calculations than the graphs built up
naively from the generators.  However, formalising spiders requires
moving from finite graphs, where each vertex has bounded degree, and
which are subject to a finite number of rewrite rules, to a system
where nodes may have arbitrarily many edges, and there are infinitely
many concrete rewrite rules. The desire to retain these intuitive
reasoning methods motivates the extension from concrete graphs to
\emph{graph patterns}. These will comprise the main subject of this
paper.

@@

graph-pattern-spider-law-fig
